Writing a reproducible paper with RStudio and Quarto

### 2023-03-25 paper.qmd
remove stargazer package + resolve overlap in latex modelsummary tables.

To name prerequisites, we assume that you are familiar with using R and R Markdown. If you don't know what R Markdown is there are many great resources that you can use (e.g. watch this [short video](https://vimeo.com/178485416)). An older template of ours [see @Bauer2018-dl; https://osf.io/q395s] may provide a quick entry point to writing a reproducible manuscript with R Markdown and Latex. To further deepen your understanding of the different components underlying this template, go on and learn about R, [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Quarto](https://quarto.org/). For Quarto also consider reading their [FAQ](https://quarto.org/docs/faq/rmarkdown.html).

@@

@tbl-4 shows a summary table for numeric variables.

``` {{r}}
#| label: tbl-4
#| tbl-cap: "Summary: Numeric variables"
library(modelsummary)
datasummary_skim(swiss, 
                 type="numeric", 
                 histogram=T,
                 output="flextable",
                 title = "Summary: Numeric variables")
```

```{r}
#| label: tbl-4
#| tbl-cap: "Summary: Numeric variables"

library(modelsummary)
datasummary_skim(swiss, 
                 type="numeric", 
                 histogram=T, 
                 output= "flextable",
                 title = "Summary: Numeric variables")
```

@@

@tbl-5 shows a summary table for categorical variables.

``` {{r}}
#| label: tbl-5
#| tbl-cap: "Summary: Categorical variables"
# Create categorical variables
swiss$Education_cat <- cut(swiss$Education, 
                   breaks=c(-Inf, 6, 12, Inf), 
                   labels=c("low","middle","high"))
swiss$Infant.Mortality_cat <- cut(swiss$Infant.Mortality, 
                   breaks=c(-Inf, 18.15, 21.70, Inf), 
                   labels=c("low","middle","high"))

@@

library(dplyr)
tab_cat <- tab_cat %>%
  font(fontname="Times New Roman", part="header") %>%
  font(fontname="Times New Roman", j=1:4) %>% 
  fontsize(size=12, part="header") %>%
  fontsize(size=10, j=1:4) %>% 
  line_spacing(space = 0.3, part = "all")

@@

```{r}
#| label: tbl-5
#| tbl-cap: "Summary: Categorical variables"

@@

library(dplyr)
tab_cat <- tab_cat %>%
  font(fontname="Times New Roman", part="header") %>%
  font(fontname="Times New Roman", j=1:4) %>%
  fontsize(size=12, part="header") %>%
  fontsize(size=12, j=1:4) %>%
  line_spacing(space = 0.3, part = "all")

@@

@tbl-7 shows the output for a regression table. Make sure you name all the models you estimate with unique identifiers (even if its 50 models) and explicitly refer to model names (M1, M2 etc.) in the text.

``` {{r}}
#| label: tbl-7
#| tbl-cap: "Linear regression with modelsummary"
library(modelsummary)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
models <- list("M1" = M1, "M2" =  M2, "M3" = M3)

@@

# Produce right output depending on format (html vs. latex)
if(knitr::is_html_output()){table7}else{table7 %>% gt::as_latex() %>% as.character() %>% gsub("\\(Intercept\\)", "Intercept", .) %>% cat()}
```

```{r}
#| label: tbl-7
#| tbl-cap: "Linear regression with modelsummary"
#| results: "asis"

library(modelsummary)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
models <- list("M1" = M1, "M2" =  M2, "M3" = M3)

# Hier Weiter

library(gt)
library(gtsummary)
# additionally we want to change the font, font size and spacing
table7 <- modelsummary(models,
             output = 'gt',
             stars = TRUE,
             notes = "Notes: put some notes here...",
             threeparttable = TRUE) %>%
    tab_spanner(label = 'Dependent variable: Fertility', columns = 2:4)

if(knitr::is_html_output()){table7}else{table7 %>% gt::as_latex() %>% as.character() %>% gsub("\\(Intercept\\)", "Intercept", .) %>% cat()}
```

>>>>

To name prerequisites, we assume that you are familiar with using R and R Markdown. If you don't know what R Markdown is there are many great resources that you can use (e.g. watch this [short video](https://vimeo.com/178485416)). An older template of ours [see @Bauer2018-dl; https://osf.io/q395s] may provide a quick entry point to writing a reproducible manuscript with R Markdown and Latex. To further deepen your understanding of the different components underlying this template, go on and learn about R, [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Quarto](https://quarto.org/). For more information on Quarto also consider reading their [FAQ](https://quarto.org/docs/faq/rmarkdown.html).

@@

@tbl-2 shows a summary table for numeric variables.

``` {{r}}
#| label: tbl-2
#| tbl-cap: "Summary: Numeric variables"
library(modelsummary)
datasummary_skim(swiss, 
                 type="numeric", 
                 histogram=F,
                 output="flextable") %>% 
  # column widths
  flextable::autofit()
```

```{r}
#| label: tbl-2
#| tbl-cap: "Summary: Numeric variables"

library(modelsummary)
datasummary_skim(swiss, 
                 type="numeric", 
                 histogram=F, 
                 output= "flextable") %>% 
  # column widths
  flextable::width(1, width = 1) %>% 
  flextable::width(2:8, width = 0.5)

```

@@

@tbl-3 shows a summary table for categorical variables. Notice how for this table we also changed the font to be "Times New Roman".

``` {{r}}
#| label: tbl-3
#| tbl-cap: "Summary: Categorical variables"
# Create categorical variables
swiss$Education_cat <- cut(swiss$Education, 
                   breaks=c(-Inf, 6, 12, Inf), 
                   labels=c("low","middle","high"))
swiss$Infant.Mortality_cat <- cut(swiss$Infant.Mortality, 
                   breaks=c(-Inf, 18.15, 21.70, Inf), 
                   labels=c("low","middle","high"))

@@

library(dplyr)
tab_cat <- tab_cat %>%
  font(fontname="Times New Roman", part="header") %>%
  font(fontname="Times New Roman", j=1:4) %>% 
  fontsize(size=12, part="header") %>%
  fontsize(size=10, j=1:4) %>% 
  line_spacing(space = 0.3, part = "all") %>% 
  # column widths
  flextable::autofit()

@@

```{r}
#| label: tbl-3
#| tbl-cap: "Summary: Categorical variables"

@@

library(dplyr)
tab_cat <- tab_cat %>%
  font(fontname="Times New Roman", part="header") %>%
  font(fontname="Times New Roman", j=1:4) %>%
  fontsize(size=12, part="header") %>%
  fontsize(size=12, j=1:4) %>%
  line_spacing(space = 0.3, part = "all") %>% 
  # column widths
  flextable::autofit()

@@

@tbl-4 shows the output for a regression table. Make sure you name all the models you estimate with unique identifiers (even if its 50 models) and explicitly refer to model names (M1, M2 etc.) in the text.

``` {{r}}
#| label: tbl-4
#| tbl-cap: "Linear regression with modelsummary"
library(modelsummary)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
models <- list("M1" = M1, "M2" =  M2, "M3" = M3)

@@

# Produce right output depending on format (html vs. latex)
if(knitr::is_html_output()){table4}else{table4 %>% gt::as_latex() %>% as.character() %>% gsub("\\(Intercept\\)", "Intercept", .) %>% cat()}
```

```{r}
#| label: tbl-4
#| tbl-cap: "Linear regression with modelsummary"
#| results: "asis"
library(modelsummary)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
models <- list("M1" = M1, "M2" =  M2, "M3" = M3)

library(gt)
library(gtsummary)
# additionally we want to change the font, font size and spacing
table4 <- modelsummary(models,
             output = 'gt',
             stars = TRUE,
             notes = "Notes: put some notes here...",
             threeparttable = TRUE) %>%
    tab_spanner(label = 'Dependent variable: Fertility', columns = 2:4)

if(knitr::is_html_output()){table4}else{table4 %>% gt::as_latex() %>% as.character() %>% gsub("\\(Intercept\\)", "Intercept", .) %>% cat()}
```
